Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-08 manuscript/manuscript.tex
ms updates

@@

>>>>

%\vb{This implies that the $N_e$
%without the effects of selection is at XXX $\nicefrac{}{(1-G(t))}$}

@@

\subsection{Approximating the Reduction in Diversity from $G(t)$}

To help help reconcile our measure of linked selection, $G(t)$, with familiar
reductions in $N_e$, or, equivalently neutral diversity, we develop some rough
approximations here. First, we introduce some convenient notation. Let $V_{T} =
\var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS} =
\sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
contribution of all pairwise temporal covariances (observable), $V_D$ be the
unobservable drift-only variance in allele frequency, and $V_{LS} =
\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i$ which is the (unobservable) effect
linked selection has on the variances in allele frequency change.

\begin{align}
  V_{T} = V_{LS} + C_{LS} + V_D
\end{align}

Our measure $G(t)$ is then,

\begin{align}
  G(t) = \frac{C_{LS}}{V_T}
\end{align}

meaning we can express the fraction of total variance due to drift alone as

\begin{align}
  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
                  &= 1 - G(t) - \varepsilon \\
                  & \le 1 - G(t).
\end{align}

where $\varepsilon \ge 0$ since these variances must be positive.

For small $t$, it is common to approximate $N_e$ as,

\begin{align}
  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
\end{align}

Then, the drift-only $N_e$ (that is, removing the effects of linked selection)
replaces $V_T$ with $V_D$,

\begin{align}
  N_{e,D} &\approx \frac{t p_0(1-p_0)}{2 V_T(1 - G(t) - \varepsilon)} \\
          &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
          &\gtrsim  \frac{N_e}{1 - G(t)}
\end{align}

In the linked selection literature, it is common to translate the impact of
linked selection as a reduction in the background level of neutral pairwise
diversity, $\pi_0$. Under the coalescent, $\pi = 2 \mu \E(T_2)$ where $\E(T_2)$
is the pairwise time to coalescence, which under drift alone in a constant
population of size $N$, is $\E(T_2) = 2N$.
